  ---
title: "Phenograph clustering"
author: "Xiaofei YU"
date: "08-1-2025"
output: 
  html_document:
    theme:
      spacelab
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
    number_sections: false
    highlight: textmate 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(plyr)
library(stringr)
library(Rphenograph)
library(dplyr)
```

# Van Maldegem IMC pipeline
 ##The most impoortant part is set makers,here i use COI_names = c( "MI_CD45", "MI_CD4", "MI_CD3e", "MI_CD8a","MI_B220","MI_Foxp3","MI_TCRgdt",
              "MI_CD68",  "MI_EpCAM", "MI_F480",  "MI_LY6G","MI_CD103", "MI_CD11b", "MI_CD11c", "MI_MHCII", "MI_Ly6C", 
          "MI_PECAM", "MI_aSMA",  "MI_CD44", "MI_CK19") 

## Set global variables

In this section, we set the input and output directories of this script and
select a data file. This file should contain normalised and scaled IMC data.
Furthermore, we create a vector with the names of the columns that should
be used for clustering.

In this data analysis strategy, we only use lineage markers for clustering to
optimise the clustering results. 

Q: Why these markers?

```{r global-variables, include=TRUE, results=FALSE}
# Project folder path:
BASE = "/mnt/Data1/groupfebe/runs/Xiaofei/CCR2_LUNG/"

# Name of input and output directories:
OUTPUT_PATH = file.path(BASE, "Data/Annotation_input")

# Name of the file wi th normalised and scaled data:
celldata =read.csv("/mnt/Data1/groupfebe/runs/Xiaofei/CCR2_LUNG/Results/Normalscale_output/250107_normalised_concatenated_scaled_data_001threshold.csv")
# check the celldata
unique(colnames(celldata))
unique(celldata$ROI_ID)
unique(celldata$MouseID)
unique(celldata$ExpGroup)
unique(celldata$domain)
unique(celldata$domain2)
unique(celldata$filename)
unique(celldata$name)


# Add a new column "name" by combining "ROI_ID" and "filename",if the post preprocessing include this step  then dont have to do it here,
celldata <- celldata %>%
  mutate(name = paste(ROI_ID, filename, sep = "_"))
unique(celldata$name)
# Names of columns to be used for clustering: 
COI_names = c( "MI_CD45", "MI_CD4", "MI_CD3e", "MI_CD8a","MI_B220","MI_Foxp3",
              "MI_CD68",  "MI_EpCAM", "MI_F480",  "MI_LY6G","MI_CD103", "MI_CD11b", "MI_CD11c", "MI_MHCII", "MI_Ly6C",
              "MI_PECAM", "MI_aSMA",  "MI_CD44", "MI_CK19")
#checking the name
COI_names1 <- COI_names[!COI_names %in% colnames(celldata)]
COI_names1
kList = c(30,25,20,15) #c(75, 50, 40, 30, 25, 20, 15, 10)
    ```

## Phenograph clustering

### Create object with clustering with multiple k's

```{r}
# Create a clustering object ---------------------------------------------------
# Description : This function creates a dataframe that holds clustering annotation
#               with different k levels
# ---Parameters---
# celldata  --> [dataframe] dataframe where rows denote individual cells, and
#               columns denote mean intensity of each channel
# COInames  --> [vector] names of the columns that should be used for clustering
# Klist     --> [vector] value between 0 and 1 that indicates the percentile
# << Output >>  [dataframe] dataframe where rows denote individual cells, and
#               columns hold cluster annotation with different k levels
set.seed(211)
create.clustering.object = function(celldata, COInames, Klist) {
  clustering = data.frame("index" = 1:nrow(celldata))
  for (k in Klist) {
    print(Sys.time())
    print(paste0("Current k: ", k))
    # Run Rphenograph on data of select columns
    Rphenograph_out <- Rphenograph(celldata[COI_names], k=k)
    
    # Create a dataframe with the annotation for each cell and sort by index
    annotation = as.vector(membership(Rphenograph_out[[2]]))
    index = as.numeric(names(membership(Rphenograph_out[[2]])))
    clusterDF = data.frame(index, annotation)[order(index),]
    
    # Find the indexes of cells that were not clustered
    allCellIndex = as.numeric(1:nrow(celldata))
    missing = allCellIndex[!allCellIndex %in% clusterDF$index]
    
    # Set the cluster for all unclustered cells to NA:
    if (length(missing) > 0) {
      unassignedCluster = NA
      
      for (n in missing) {
        clusterDF = rbind(clusterDF[1:(n-1),],
                          c(n, unassignedCluster),
                          clusterDF[n:nrow(clusterDF),])
      }
    }
    
    # Add clustering to the celldata dataframe
    if (isTRUE(all.equal(clustering$index, clusterDF$index))){
      colnames = c(colnames(clustering), paste0("k", k))
      clustering = cbind(clustering, as.numeric(clusterDF$annotation))
      colnames(clustering) = colnames
    } else {
      stop("Indexes do not correspond")
    }
  }
  
  clustering = subset(clustering, select = -c(index))
  return(clustering)
}

clusteringObject = create.clustering.object(celldata, COI_names, kList)

# Save data with clustering information added
date = str_sub(gsub("-", "", Sys.Date()), -6, -1)
file_path = paste0(OUTPUT_PATH, "/", date, "_clustering_k",
                   paste(kList, collapse = "_k"),
                   "_", length(COI_names), "markers.csv")
write.csv(clusteringObject, file = file_path, row.names = F)
print(Sys.time())
```


